---
title: "Project 2"
author: "Elisa Rwagasana Ishimwe"
date: "2024-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "OneDrive - The College of Wooster/Desktop/Applied DS")
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{r}
#Importing the base dataset
library(readr)
train_base <- read_csv("~/Desktop/Project 2/home-credit-credit-risk-model-stability/csv_files/train/train_base.csv")
```
```{r}
#Importing train person_1
library(readr)
train_person_1 <- read_csv("~/Desktop/Project 2/home-credit-credit-risk-model-stability/csv_files/train/train_person_1.csv")
```


```{r}
#Merging train_person_1 to the base table
# Outer join
data <- merge(train_base, train_person_1, by = "case_id", all = TRUE)
```

```{r}
#Data manipulation 
#Selecting the variables needed 
data <- data %>%
  dplyr::select(
    case_id, 
    mainoccupationinc_384A, 
    birth_259D,
    target,
    familystate_447L,
    sex_738L,
    incometype_1044T
  )
```


```{r}
#Remove NAs
data <- data %>%
  filter(
    !is.na(mainoccupationinc_384A) & 
    !is.na(birth_259D) & 
    !is.na(familystate_447L) & 
    !is.na(sex_738L)
  )%>%
  sample_n(size = 100000, replace = FALSE)
```


```{r}
#Change my date variable 
data <- data %>% mutate (target = as.factor(target))%>%
  separate(col = birth_259D, into = c("year", "month", "day"), sep = "-")
```

```{r}
library(dplyr)

# Assuming 'data' is your data frame and 'year' is the variable representing the year
data <- data %>%
  mutate(year = as.numeric(year),  # Convert 'year' to numeric
         generation = ifelse(between(year, 1925, 1945), "Silent Generation",
                              ifelse(between(year, 1946, 1964), "Baby Boomers",
                                     ifelse(between(year, 1965, 1980), "Generation X",
                                            ifelse(between(year, 1981, 1996), "Millennials",
                                                   ifelse(between(year, 1997, 2012), "Generation Z", "Other"))))))



```

```{r}
#Saving dataset
write.csv(data, "project2data.csv")
```


```{r}
str(data)
```
```{r}
#Splitting into train and test
set.seed(1)
n <- nrow(data)

# Generate random indices for the training set (80%)
train_indices <- sample(1:n, size = 0.8 * n, replace = FALSE)

# Create the training set
train_set <- data[train_indices, ]

# Create the test set by excluding the training set
test_set <- data[-train_indices, ]
```


```{r}
#Model building 
glm.fit1 <- glm(target ~ generation + familystate_447L + sex_738L + incometype_1044T
               , data = train_set
               , family = binomial)
summary(glm.fit1)
```


# Assuming merged_dataset is your dataframe
library(rpart)
library(rpart.plot)

# Create a regression or classification tree
tree_model <- rpart(target ~ mainoccupationinc_384A + year + familystate_447L + sex_738L,
                    data = tr,
                    method = "class",  # Use "class" for classification, "anova" for regression
                    cp = 0.01 
)

# Display the tree
print(tree_model)
rpart.plot(tree_model)

```{r}
# Assuming glm.fit1 is your initial model
library(caret)
library(lattice)
library(MASS)
backward_model <- stepAIC(glm.fit1, direction="backward")
summary(backward_model)

```
Our model with all predictors is the best because it has a lower AIC and smaller deviance.

# Classification tree (NOT WORKING)
```{r}
# Fit the classification tree with specified cp
class_tree <- rpart(target ~.,
                    data = train_set,
                    method = "class",
                    control = rpart.control(cp = 0.001))  

# Load the rpart.plot library
library(rpart.plot)

# Plot the classification tree
rpart.plot(class_tree)
```

# Random Forest
```{r}
library(randomForest)
rf_model <- randomForest(target ~ mainoccupationinc_384A + generation + incometype_1044T +  
    familystate_447L + sex_738L
                  , data = train_set
                  , mtry = 7
                  , ntree = 300
                  , importance = TRUE
                  , type = "class")
rf_model
varImpPlot(rf_model)
```


```{r}
# Predicted probabilities
predicted_probabilities <- predict(glm.fit1, newdata = test_set, type = "response")

# Convert probabilities to binary predictions using a threshold (e.g., 0.5)
class_pred <- ifelse(predicted_probabilities > 0.5, 1, 0)
```

```{r}
# Compare predicted values with the actual response variable in the test set
conf_matrix <- table(Actual = test_set$target, Predicted = class_pred)

# Display the confusion matrix
print(conf_matrix)

# Calculate accuracy
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
cat("Accuracy:", accuracy, "\n")

```
True Negatives (TN): 19351 (actual negative instances correctly predicted as negative)
False Negatives (FN): 649 (actual positive instances incorrectly predicted as negative)
It indicates that your model is correctly identifying the majority of negative instances (TN) but is missing a significant portion of positive instances (FN).
Weakness: This suggests that your model is conservative in predicting positive instances and tends to under predict them.
It is only working when the threshold is equal to 0.1 (which is too low). There might also be because of class imbalance where there is a balance in the distribution of positive and negative instances in the dataset. 

### EDA
```{r}
#Target with main income
data %>%
  ggplot(aes(x = target, y = mainoccupationinc_384A)) +
  geom_boxplot()+
  labs(title = "Default status by Amount of Income",
       x = "Default Status",
       y = "Amount of income")
  
```


```{r}
# Assuming 'count' is the variable representing the count of occurrences
data %>%
  group_by(generation, target) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(generation,count), y = count, fill = target)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Default Status by Generation",
       x = "Generation",
       y = "Count",
       fill = "Target")

library(ggplot2)

data %>%
  filter(target == 1) %>%
  group_by(generation) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(generation, count), y = count, fill = generation)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Defaulting on a Loan Across Generations",
       x = "Generation",
       y = "Number of People",
       fill = "Generation") + 
  scale_fill_brewer(palette = "Oranges") +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),  # Rotate x-axis labels vertically
    panel.grid.major = element_blank(),      # Remove major grid lines
    panel.grid.minor = element_blank(),      # Remove minor grid lines
    panel.background = element_blank(),      # Remove panel background
    axis.line = element_line(color = "black"),  # Add axis lines
    axis.ticks = element_blank()             # Remove axis ticks
  )

```

```{r}
library(ggplot2)

data %>%
  filter(target == 1) %>%
  group_by(familystate_447L) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(familystate_447L, count), y = count, fill = familystate_447L)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Defaulting on a Loan Across Marital Status",
       x = "Marital Status",
       y = "Number of People",
       fill = "Marital Status") + 
  scale_fill_brewer(palette = "Reds") +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),  # Rotate x-axis labels vertically
    panel.grid.major = element_blank(),      # Remove major grid lines
    panel.grid.minor = element_blank(),      # Remove minor grid lines
    panel.background = element_blank(),      # Remove panel background
    axis.line = element_line(color = "black"),  # Add axis lines
    axis.ticks = element_blank()             # Remove axis ticks
  )

```

```{r}
#Target with gender
data %>%
  group_by(sex_738L, target) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = sex_738L, y = count, fill = target)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Barplot of Two Categorical Variables",
       x = "Year",
       y = "Count",
       fill = "Target")
```

```{r}
# Create a contingency table
contingency_table <- table(data$target, data$familystate_447L)

# Calculate proportions
proportions <- prop.table(contingency_table, margin = 1)  # Proportions within each row (category1)
# Alternatively, you can calculate proportions within each column (category2) by changing margin = 2

# Print the proportions
print(proportions)

```
```{r}
data %>%
  filter(target == 1, incometype_1044T %in% c("SELFEMPLOYED", "RETIRED_PENSIONER", "SALARIED_GOVT", "EMPLOYED", "PRIVATE_SECTOR_EMPLOYEE")) %>%
  group_by(incometype_1044T) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(incometype_1044T, count), y = count, fill = incometype_1044T)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Defaulting on a Loan Across Different Income Types",
       x = "Income Type",
       y = "Number of People") + 
  scale_fill_manual(values = c("PRIVATE_SECTOR_EMPLOYEE" = "darkred", # Highlight PRIVATE_SECTOR_EMPLOYEE in red
                               "Other" = "grey")) +           # Other categories in grey
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),  # Rotate x-axis labels vertically
    panel.grid.major = element_blank(),      # Remove major grid lines
    panel.grid.minor = element_blank(),      # Remove minor grid lines
    panel.background = element_blank(),      # Remove panel background
    axis.line = element_line(color = "black"),  # Add axis lines
    axis.ticks = element_blank()             # Remove axis ticks
  )

```

